---
title: "Analysis of the results of the US presidential election in 2020"
---

# Report purpose and data source

The purpose of the report is to analyze the results of the US presidential election in 2020 at the level of individual states and appropriate visualization using an interactive map.

Source: <https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/42MVDX>

# Visualization

```{r message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(maps)
library(mapdata)
library(dplyr)
library(tidyr)
library(plotly)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
election <-
  read.csv("1976-2020-president.csv",sep = ",")
election <-
  election %>% mutate(result_in_percent = round(candidatevotes / totalvotes *
                                                  100, 2),party_detailed= gsub("DEMOCRATIC-FARMER-LABOR", "DEMOCRAT", party_detailed)) %>% filter(state != "DISTRICT OF COLUMBIA")
election2 <-
  election %>% filter(party_detailed == "DEMOCRAT" |
                        party_detailed == "REPUBLICAN") %>% select(c(1, 2, 8, 9))
election2 <-
  pivot_wider(election2, names_from = party_detailed, values_from = candidate,values_fn = list(candidate = first))
election3 <-
  election %>% filter(party_simplified == "DEMOCRAT" |
                        party_detailed == "REPUBLICAN") %>% select(c(1, 2, 15, 16))
election3 <-
  pivot_wider(election3, names_from = party_simplified, values_from = result_in_percent,values_fn = list(result_in_percent = first))

election3<-election3 %>% mutate(DEMOCRAT_result=DEMOCRAT,REPUBLICAN_result=REPUBLICAN)

election_final<-cbind(election2,election3[,c(5,6)])

election_final <-
  election_final %>% mutate(
    DEMOCRAT = gsub(",", "", DEMOCRAT),
    REPUBLICAN = gsub(",", "", REPUBLICAN),
    OTHERS_result = 100 - unlist(DEMOCRAT_result) - unlist(REPUBLICAN_result)
  )
winners <-
  election %>% group_by(year, state) %>%
  slice(which.max(result_in_percent)) %>%
  select(c("year", "state", "party_simplified"))

election_final$winner<-winners$party_simplified

    g <- list(
      scope = 'usa',
      projection = list(type = 'albers usa'),
      lakecolor = toRGB('white')
    )
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
election_final2 <-
      election_final %>% filter(year == 2020)

plot_geo() %>%
      add_trace(
        fillcolor = ~ election_final2$winner,
        span = I(0),
        locations = state.abb,
        text = paste0(
          "State: ",
          election_final2$state,
          "<br>",
          election_final2$DEMOCRAT,
          " (DEMOCRAT): ",
          election_final2$DEMOCRAT_result,
          "%",
          "<br>",
          election_final2$REPUBLICAN,
          " (REPUBLICAN): ",
          election_final2$REPUBLICAN_result,
          "%",
          "<br>",
          "OTHERS: ",
          round(election_final2$OTHERS_result,2),
          "%"
        ),
        locationmode = 'USA-states',
        hovertemplate = paste('%{text}')
      ) %>%
      layout(geo = g,legend=list(title=list(text="Winner:")),hoverlabel = list(font = list(size = 9)),title ="Results of the US presidential election in 2020")

```

The map allows you to see which party won in a specific state in the US presidential election. When you hover over a specific state, it allows you to see more accurate results in a given state. In addition clicking on a party's name included in the legend will result in only the states in which it won remaining selected.

# Conclusions

-   We can see that the US has a two-party system because no party other than Democrats and Republicans received more than 5% in any state.
-   Analysis of the data shows that the northwestern states and Hawaii favor Democrats more, while the southern and central states and Alaska favor Republicans more.
